---
title: "Module 14"
format: html
editor: visual
---

## Confidence Intervals 

```{r}
library(boot)
library(tidyverse)
library(manipulate)
library(mosaic)

```

### Confidence intervals 

is another way of describing a statistic’s sampling distribution, and it plays a central role in basic inferential statistics.

the confidence interval is an interval around our estimate of mean of the sampling distribution for a particular statistic (typically a mean), and it gives us a range of values into which subsequent estimates of a statistic would be expected to fall some critical proportion of the time, if the sampling exercise were to be repeated.

The “95% CI” around a statistic, for example, describes the range of values into which a new estimate of the statistic, derived from a subsequent sample, would be expected to fall 95% of the time. The “99% CI” around the initial statistic would be slightly wider, while the “50% CI” would be narrower.

### Calculating Theoreticals CIs

For large sample sizes ( n ≥ \~30) = use the normal distribution for ur theoretical distribution used to define the critical values

Smaller sample sizes ( n \< \~30), = *Student’s t distribution*

### Interpretation of CIs

-   Based on the given data (with a particular mean, variance, and sample size), we are X% confident that the true mean of the population lies between the lower and upper bounds.

-   The mean of a repeated sample of the same size drawn from this same underlying distribution is expected to fall into the given interval X% of the time.

## The Central Limit Therom

random variables approaches a normal distribution as the sample size increases. It is this fact that allows us to have a good sense of the mean and distribution of average events in a population even though we only observe one or a small number of samples of those events and even though we do not know what the actual population distribution is!

```{r}
lambda <- 14
n <- 1000
(se <- sqrt(lambda/n))

```

```{r}
# the estimated SE, here, based on an estimate of the population variance if
# available
par(mfrow = c(1, 3))
hist(rpois(10000, lambda = lambda), probability = TRUE, xlab = "x", main = "Original Distribution")
x <- vector(length = 1000)
for (i in 1:1000) {
    x[[i]] <- mean(rpois(n = n, lambda = lambda))
}
hist(x, breaks = seq(from = min(x), to = max(x), length.out = 20), probability = TRUE,
    xlab = "Mean X", main = "Sampling Distribution")
(se <- sd(x))
```

```{r}
# the estimated SE, here based on the standard deviation of the actual sampling
# distribution
qqnorm(x)
qqline(x)
```
