---
title: "Module 11 functions and flow"
format: html
editor: visual
---

## Module 11 Functions and flow control

```         
```

```{r}
library(sjmisc)
library(tidyverse)
```

Flow control if something then do else

```{r}
i <- TRUE

if (i == TRUE) {
    print("Yes")
} else {
    print("No")
}

```

ifelse() function \<--- There is also the if_else function are teacher uses much more.

```{r}

i <- 9
ifelse(i <= 10, "Yes", "No")

```

For mixing several "if... else..." statemtns use --\> case_when

```{r}
i <- 1:10
output <- case_when(i <= 3 ~ "small", i <= 7 ~ "medium", i <= 10 ~ "large")
output
```

If you have a list of numbers these fucntions will tell you a little about all of them

```{r}
i<- c(9,10,11)
any(i <= 10)
all(i<=10)
```

## relational operators

-   less than, greater than: `<`, `>`

-   less than or equal to, greater than or equal to: `<=`, `>=`

-   equal to: `==` **NOTE:** This uses a double equal sign!

-   not equal to: `!=`

## Iterating with Loops
for()
example
```{r}
v<-seq(from=100, to= 120, by = 2)
for (i in 1: length(v)) {
  #here, we are looping ober the indices of v 
  print(v[i])
}
```
